Data Preparation

### 2021-06-29 x0_data-preparation.Rmd
included thresholds graph and formulas

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(zoo)

@@

```{r, dpi=480}
#THRESHOLD AT ROLLING QUANTILES PER YEAR
#Please install the "roll" package.
#install.packages("roll")
library(roll)

ds1 <- full_dataset %>% group_by(YR = year(DATE), DATE) %>% summarize(N = n(), .groups = "drop") %>% mutate(THRESHOLD = quantile(N, probs = 0.7))
ds1

ds1 %>% ggplot(aes(x = DATE)) +
  geom_line(aes(y = THRESHOLD)) +
  geom_point(aes(y = N, color = N > THRESHOLD))

#CHECK WEEKLY DISTRIBUTION
ds2 <- ds1 %>% mutate(WEEK_DAY = wday(DATE, label = TRUE), W_END = WEEK_DAY %in% c("Sat", "Sun"))
uplim <- .95
lowlim <- .20

ds2 <- ds2 %>% mutate(MED = roll_quantile(N, width = 7, p = .50, min_obs = 7),
                      UP_THR = roll_quantile(N, width = 180, p = uplim, min_obs = 7),
                      LOW_THR = roll_quantile(N, width = 180, p = lowlim, min_obs = 7))
ds2

ds2 %>% ggplot(aes(x = DATE)) +
  geom_col(aes(y = MED, fill = !(MED < LOW_THR | MED > UP_THR))) +
  geom_line(aes(y = UP_THR), color = "blue") +
  geom_line(aes(y = LOW_THR), color = "red") +
  scale_x_date(limits = c(as.Date("2019-10-01"), as.Date("2020-12-01")), date_breaks = "30 days", date_labels = "%b\n%y") +
  labs(y = "7-day rolling median", x = NULL, fill = "Normal operation?",
       title = "What is normal?",
       subtitle = "Red line - Lower threshold; Blue line - upper threshold",
       caption = "Lower threshold set to 20th percentile of the last 180 days;\nUpper threshold set to 95th percentile of the last 180 days"
)

#ds2 %>% ggplot() +
  geom_freqpoly(aes(x = N, color = W_END))
```
